Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2019-10-22 paper2_JNLS/paper.tex
NM

	Now, from the perspective of statistical mechanics, consider the communication between the statistical ensembles, $\Gupdn$, upstream and downstream of the ADC. These two systems are in contact at the ADC, and so the upstream state $\Gup$ can be regarded as a thermal reservoir that influences the downstream state $\Gdn$. In fact, the above deterministic  condition directly leads to a simple description for the link. The quantity of interest is the outgoing Hamiltonian $\Hdn$, since its value just upstream of the ADC, $t=0^{-}$, is set by the incoming dynamics, and then this particular value is conserved thereafter in the outgoing dynamics. Since $u$ matches at the ADC, so must $\Hdn$, and, in fact, this matching holds for every individual trajectory. Recall that $\Hdn(t)$ is not conserved in the upstream dynamics and so its value varies for $t < 0$. However, appealing to weak Ergodicity, the ensemble mean $\meanup{\Hdn(t)}$ is expected to be independent of time. In particular, the value $\meanup{\Hdn \vert_{t=0^{-}} }$ is the same as the bare ensemble mean $\meanup{\Hdn}$. Afterwards, the particular value $\Hdn \vert_{t=0^{-}}$ is conserved in the downstream dynamics, on a trajectory-by-trajectory basis and for all $t>0$. Thus, the downstream measure $\Gdn$ must recover the same ensemble mean $\meanup{\Hdn \vert_{t=0^{-}} } = \meanup{\Hdn}$, producing the simple condition
\begin{align}
\label{statmatch}
\meanup{\Hdn} = \meandn{\Hdn}
\qquad \mbox{\em Statistical matching condition}
\end{align}

@@

	The invariant measure exhibits the proportionality $ \Gibbs_{\Lambda} \propto \exp(-\theta_{\Lambda} \Ham_{\Lambda})$, where we have made explicit the dependence of all quantities on $\Lambda$. In particular, if $\Ham_{\Lambda}$ where to depend sensitively on $\Lambda$ in expectation, then $\theta_{\Lambda}$ would need to compensate in order to produce the same invariant measure. Hence, it would be desirable to scale the system in such a way that $\Ham_{\Lambda}$ does not depend sensitively on $\Lambda$, at least in expectation. 
To achieve this insensitivity, we will appeal to the uniform measure $\Gz$ and the idea of equipartition of energy \cite{abramov2003}, since these two concepts afford simple scaling estimates.

Recall that $\Ham_{\Lambda}$ is composed of the cubic and quadratic components $\Hthree$ and $\Htwo$. Due to the odd symmetry of $\Hthree$, it is easy to see that $\meanz{\Hthree} = 0$. The quadratic component, however, requires closer inspection. Due to Parseval's identity, $\Htwo$ can be written as
\begin{equation}
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx = 2 \pi \sum_{k=1}^{\Lambda} k^2 \abs{\uhat_k}^2
\end{equation}
Due to the constraint $\En[\uL] = 1$, the equipartioned microstate is given by 
\begin{align}
&\abs{\uhat_k}^2 \approx \frac{1}{2 \pi \Lambda}	 \qquad \mbox{\em for equipartition of energy}
\end{align}
Then the expected value of $\Htwo$ under the uniform measure is
\begin{equation}
\label{HtwoExpect}
\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} {k}^ 2 \mean{\abs{\uhat_k}^2}_0 \sim \frac{1}{3} \Lambda^2
\end{equation}
where we have used the identity for the Gauss-like sum
\begin{equation}
\sum_{k=1}^{n} {k}^ 2 = \frac{1}{6} n(n+1)(2n+1) \approx \frac{1}{3} n^3
\end{equation}

>>>>

	Now, from the perspective of statistical mechanics, consider the communication between the statistical ensembles, $\Gupdn$, upstream and downstream of the ADC. These two systems are in contact at the ADC, and so the upstream state with distribution $\Gup$ can be regarded as a thermal reservoir that influences the downstream distribution $\Gdn$. In fact, the above deterministic  condition directly leads to a simple description for the link. The quantity of interest is the outgoing Hamiltonian $\Hdn$, since its value just upstream of the ADC, $t=0^{-}$, is set by the incoming dynamics, and then this particular value is conserved thereafter in the outgoing dynamics. Since $u$ matches at the ADC, so must $\Hdn$, and, in fact, this matching holds for every individual trajectory. Recall that $\Hdn(t)$ is not conserved in the upstream dynamics and so its value varies for $t < 0$. However, appealing to weak Ergodicity, the ensemble mean $\meanup{\Hdn(t)}$ is expected to be independent of time. In particular, the value $\meanup{\Hdn \vert_{t=0^{-}} }$ is the same as the bare ensemble mean $\meanup{\Hdn}$. Afterwards, the particular value $\Hdn \vert_{t=0^{-}}$ is conserved in the downstream dynamics, on a trajectory-by-trajectory basis and for all $t>0$. Thus, the downstream measure $\Gdn$ must recover the same ensemble mean $\meanup{\Hdn \vert_{t=0^{-}} } = \meanup{\Hdn}$, producing the simple condition
\begin{align}
\label{statmatch}
\meanup{\Hdn} = \meandn{\Hdn}
\qquad \mbox{\em Statistical matching condition}
\end{align}

@@

	The invariant measure exhibits the proportionality $ \Gibbs_{\Lambda} \propto \exp(-\theta_{\Lambda} \Ham_{\Lambda})$, where we have made explicit the dependence of all quantities on $\Lambda$. In particular, if $\Ham_{\Lambda}$ were to depend sensitively on $\Lambda$ in expectation, then $\theta_{\Lambda}$ would need to compensate in order to produce the same invariant measure. Hence, it would be desirable to scale the system in such a way that $\Ham_{\Lambda}$ does not depend sensitively on $\Lambda$, at least in expectation. 
To achieve this insensitivity, we will appeal to the uniform measure $\Gz$ and the idea of equipartition of energy \cite{abramov2003}, since these two concepts afford simple scaling estimates.

Recall that $\Ham_{\Lambda}$ is composed of the cubic and quadratic components $\Hthree$ and $\Htwo$. Due to the odd symmetry of $\Hthree$, it is easy to see that $\meanz{\Hthree} = 0$ with respect to the uniform measure $\Gz$. The quadratic component, however, requires closer inspection. Due to Parseval's identity, $\Htwo$ can be written as
\begin{equation}
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx = 2 \pi \sum_{k=1}^{\Lambda} k^2 \abs{\uhat_k}^2
\end{equation}
Due to the constraint $\En[\uL] = 1$, the equipartioned microstate is given by 
\begin{align}
&\abs{\uhat_k}^2 \approx \frac{1}{2 \pi \Lambda}	 \qquad \mbox{\em for equipartition of energy}
\end{align}
Then the expected value of $\Htwo$ under the uniform measure is
\begin{equation}
\label{HtwoExpect}
\mean{\Htwo}_0 = 2 \pi \sum_{k=1}^{\Lambda} {k}^ 2 \mean{\abs{\uhat_k}^2}_0 \sim \frac{1}{3} \Lambda^2
\end{equation}
where we have used the identity for the Gauss-like sum
\begin{equation}
\sum_{k=1}^{n} {k}^ 2 = \frac{1}{6} n(n+1)(2n+1) \approx \frac{1}{3} n^3
\end{equation}
